---
title: "R Performance Analysis Dashboard"
author: "Kamil Nazaruk"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: darkly
    toc: true
    toc_float: true
    code_folding: hide
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 6
)

# Load required libraries
library(ggplot2)
library(dplyr)
library(plotly)
library(DT)
library(knitr)
library(scales)
```

```{css, echo=FALSE}
body {
  background-color: #0f172a;
  color: #f1f5f9;
}

.main-container {
  max-width: 1400px;
  margin-left: auto;
  margin-right: auto;
}

h1, h2, h3, h4 {
  color: #38bdf8;
}

.card {
  background-color: #1e293b;
  border-left: 4px solid #38bdf8;
  padding: 15px;
  margin: 15px 0;
  border-radius: 5px;
}

.metric-box {
  background: linear-gradient(135deg, #1e293b 0%, #334155 100%);
  border: 1px solid #38bdf8;
  border-radius: 8px;
  padding: 20px;
  text-align: center;
  margin: 10px;
}

.metric-value {
  font-size: 2em;
  font-weight: bold;
  color: #38bdf8;
}

.metric-label {
  font-size: 0.9em;
  color: #94a3b8;
  margin-top: 5px;
}
```

<div class="card">
<h1>📊 R Performance Analysis Dashboard</h1>
<p><strong>Real-time operational performance monitoring and analysis</strong></p>
<p>This interactive report analyzes system performance metrics including response times, throughput, error rates, and resource utilization.</p>
</div>

```{r generate_data, include=FALSE}
# Generate sample performance data
set.seed(42)
n_records <- 1000

performance_data <- data.frame(
  timestamp = seq(from = as.POSIXct("2024-01-01 00:00:00"),
                  by = "hour",
                  length.out = n_records),
  operation_id = paste0("OP", sprintf("%04d", 1:n_records)),
  response_time_ms = rlnorm(n_records, meanlog = 4.5, sdlog = 0.8),
  throughput_ops = rpois(n_records, lambda = 450),
  error_rate = rbeta(n_records, shape1 = 1, shape2 = 50),
  cpu_usage = rbeta(n_records, shape1 = 5, shape2 = 3),
  memory_usage_mb = rnorm(n_records, mean = 2048, sd = 512),
  success_rate = rbeta(n_records, shape1 = 95, shape2 = 5),
  latency_p95_ms = rlnorm(n_records, meanlog = 5.2, sdlog = 0.6),
  concurrent_users = rpois(n_records, lambda = 75)
)

# Add some anomalies (10% of data)
anomaly_indices <- sample(1:n_records, size = n_records * 0.10)
performance_data$response_time_ms[anomaly_indices] <-
  performance_data$response_time_ms[anomaly_indices] * runif(length(anomaly_indices), 3, 8)
performance_data$error_rate[anomaly_indices] <-
  performance_data$error_rate[anomaly_indices] + runif(length(anomaly_indices), 0.1, 0.3)

# Add derived columns
performance_data$hour <- as.POSIXlt(performance_data$timestamp)$hour
performance_data$day <- as.Date(performance_data$timestamp)
performance_data$is_anomaly <- 1:n_records %in% anomaly_indices
```

## Executive Summary

```{r summary_metrics, echo=FALSE}
# Calculate summary statistics
avg_response_time <- mean(performance_data$response_time_ms)
avg_throughput <- mean(performance_data$throughput_ops)
avg_error_rate <- mean(performance_data$error_rate) * 100
avg_cpu <- mean(performance_data$cpu_usage) * 100
anomaly_count <- sum(performance_data$is_anomaly)
anomaly_pct <- (anomaly_count / n_records) * 100
```

<div style="display: flex; flex-wrap: wrap; justify-content: space-around;">
  <div class="metric-box">
    <div class="metric-value">`r format(round(avg_response_time, 1), big.mark=",")`ms</div>
    <div class="metric-label">Avg Response Time</div>
  </div>
  <div class="metric-box">
    <div class="metric-value">`r format(round(avg_throughput, 0), big.mark=",")`</div>
    <div class="metric-label">Avg Throughput (ops/sec)</div>
  </div>
  <div class="metric-box">
    <div class="metric-value">`r sprintf("%.2f%%", avg_error_rate)`</div>
    <div class="metric-label">Avg Error Rate</div>
  </div>
  <div class="metric-box">
    <div class="metric-value">`r sprintf("%.1f%%", avg_cpu)`</div>
    <div class="metric-label">Avg CPU Usage</div>
  </div>
  <div class="metric-box">
    <div class="metric-value">`r anomaly_count`</div>
    <div class="metric-label">Anomalies Detected (`r sprintf("%.1f%%", anomaly_pct)`)</div>
  </div>
</div>

---

## Performance Trends Over Time

### Response Time Analysis

```{r response_time_trend, echo=FALSE}
# Daily aggregation
daily_stats <- performance_data %>%
  group_by(day) %>%
  summarise(
    avg_response = mean(response_time_ms),
    p50_response = median(response_time_ms),
    p95_response = quantile(response_time_ms, 0.95),
    p99_response = quantile(response_time_ms, 0.99)
  )

# Create interactive plot
p <- plot_ly(daily_stats) %>%
  add_trace(x = ~day, y = ~avg_response, type = 'scatter', mode = 'lines+markers',
            name = 'Average', line = list(color = '#38bdf8', width = 3)) %>%
  add_trace(x = ~day, y = ~p95_response, type = 'scatter', mode = 'lines',
            name = 'P95', line = list(color = '#f59e0b', dash = 'dash')) %>%
  add_trace(x = ~day, y = ~p99_response, type = 'scatter', mode = 'lines',
            name = 'P99', line = list(color = '#ef4444', dash = 'dot')) %>%
  layout(
    title = "Response Time Trends",
    xaxis = list(title = "Date", color = '#f1f5f9'),
    yaxis = list(title = "Response Time (ms)", color = '#f1f5f9'),
    plot_bgcolor = '#1e293b',
    paper_bgcolor = '#0f172a',
    font = list(color = '#f1f5f9'),
    hovermode = 'x unified'
  )

p
```

### Throughput and Error Rate

```{r throughput_errors, echo=FALSE, fig.height=5}
daily_perf <- performance_data %>%
  group_by(day) %>%
  summarise(
    avg_throughput = mean(throughput_ops),
    avg_error_rate = mean(error_rate) * 100
  )

# Throughput plot
p1 <- plot_ly(daily_perf) %>%
  add_trace(x = ~day, y = ~avg_throughput, type = 'bar',
            name = 'Throughput', marker = list(color = '#3b82f6')) %>%
  layout(
    title = "Average Throughput (ops/sec)",
    xaxis = list(title = "Date", color = '#f1f5f9'),
    yaxis = list(title = "Operations/sec", color = '#f1f5f9'),
    plot_bgcolor = '#1e293b',
    paper_bgcolor = '#0f172a',
    font = list(color = '#f1f5f9')
  )

# Error rate plot
p2 <- plot_ly(daily_perf) %>%
  add_trace(x = ~day, y = ~avg_error_rate, type = 'scatter', mode = 'lines+markers',
            name = 'Error Rate', line = list(color = '#ef4444', width = 3),
            marker = list(size = 8)) %>%
  layout(
    title = "Average Error Rate (%)",
    xaxis = list(title = "Date", color = '#f1f5f9'),
    yaxis = list(title = "Error Rate (%)", color = '#f1f5f9'),
    plot_bgcolor = '#1e293b',
    paper_bgcolor = '#0f172a',
    font = list(color = '#f1f5f9')
  )

subplot(p1, p2, nrows = 2, shareX = TRUE)
```

---

## Resource Utilization

### CPU and Memory Usage

```{r resource_usage, echo=FALSE}
# CPU usage over time
p_cpu <- plot_ly(performance_data) %>%
  add_trace(x = ~timestamp, y = ~cpu_usage * 100, type = 'scatter', mode = 'lines',
            name = 'CPU Usage', line = list(color = '#8b5cf6'),
            fill = 'tozeroy', fillcolor = 'rgba(139, 92, 246, 0.3)') %>%
  layout(
    title = "CPU Usage Over Time",
    xaxis = list(title = "Timestamp", color = '#f1f5f9'),
    yaxis = list(title = "CPU Usage (%)", color = '#f1f5f9', range = c(0, 100)),
    plot_bgcolor = '#1e293b',
    paper_bgcolor = '#0f172a',
    font = list(color = '#f1f5f9')
  )

# Memory usage distribution
p_mem <- plot_ly(performance_data, x = ~memory_usage_mb, type = 'histogram',
                 marker = list(color = '#06b6d4')) %>%
  layout(
    title = "Memory Usage Distribution",
    xaxis = list(title = "Memory Usage (MB)", color = '#f1f5f9'),
    yaxis = list(title = "Frequency", color = '#f1f5f9'),
    plot_bgcolor = '#1e293b',
    paper_bgcolor = '#0f172a',
    font = list(color = '#f1f5f9')
  )

subplot(p_cpu, p_mem, nrows = 1)
```

---

## Hourly Performance Patterns

```{r hourly_patterns, echo=FALSE}
hourly_stats <- performance_data %>%
  group_by(hour) %>%
  summarise(
    avg_response_time = mean(response_time_ms),
    avg_throughput = mean(throughput_ops),
    avg_error_rate = mean(error_rate) * 100,
    n_operations = n()
  )

# Heatmap style visualization
p <- plot_ly(hourly_stats) %>%
  add_trace(x = ~hour, y = ~avg_response_time, type = 'bar',
            name = 'Response Time (ms)', yaxis = 'y', marker = list(color = '#38bdf8')) %>%
  add_trace(x = ~hour, y = ~avg_error_rate, type = 'scatter', mode = 'lines+markers',
            name = 'Error Rate (%)', yaxis = 'y2', line = list(color = '#ef4444', width = 3),
            marker = list(size = 8)) %>%
  layout(
    title = "Hourly Performance Patterns",
    xaxis = list(title = "Hour of Day (0-23)", color = '#f1f5f9'),
    yaxis = list(title = "Response Time (ms)", color = '#38bdf8'),
    yaxis2 = list(title = "Error Rate (%)", overlaying = 'y', side = 'right', color = '#ef4444'),
    plot_bgcolor = '#1e293b',
    paper_bgcolor = '#0f172a',
    font = list(color = '#f1f5f9'),
    hovermode = 'x unified',
    legend = list(orientation = 'h', y = -0.2)
  )

p
```

<div class="card">
**Key Findings:**

- **Best Performance Hour:** `r hourly_stats$hour[which.min(hourly_stats$avg_response_time)]`:00 (lowest response time: `r round(min(hourly_stats$avg_response_time), 1)` ms)
- **Worst Performance Hour:** `r hourly_stats$hour[which.max(hourly_stats$avg_response_time)]`:00 (highest response time: `r round(max(hourly_stats$avg_response_time), 1)` ms)
- **Peak Error Hour:** `r hourly_stats$hour[which.max(hourly_stats$avg_error_rate)]`:00 (error rate: `r sprintf("%.2f%%", max(hourly_stats$avg_error_rate))`)
</div>

---

## Anomaly Detection

```{r anomalies, echo=FALSE}
# Get anomalous operations
anomalies <- performance_data %>%
  filter(is_anomaly) %>%
  select(timestamp, operation_id, response_time_ms, error_rate, cpu_usage) %>%
  arrange(desc(response_time_ms)) %>%
  head(20)

# Scatter plot of response time vs throughput
p <- plot_ly(performance_data, x = ~throughput_ops, y = ~response_time_ms,
             color = ~is_anomaly, colors = c('#22c55e', '#ef4444'),
             type = 'scatter', mode = 'markers',
             marker = list(size = 8, opacity = 0.6),
             text = ~paste('Operation:', operation_id,
                          '<br>Response Time:', round(response_time_ms, 1), 'ms',
                          '<br>Throughput:', throughput_ops, 'ops/sec',
                          '<br>Error Rate:', sprintf("%.2f%%", error_rate * 100)),
             hoverinfo = 'text') %>%
  layout(
    title = "Response Time vs Throughput (Anomalies Highlighted)",
    xaxis = list(title = "Throughput (ops/sec)", color = '#f1f5f9'),
    yaxis = list(title = "Response Time (ms)", type = 'log', color = '#f1f5f9'),
    plot_bgcolor = '#1e293b',
    paper_bgcolor = '#0f172a',
    font = list(color = '#f1f5f9'),
    showlegend = TRUE
  )

p
```

### Top 20 Anomalous Operations

```{r anomaly_table, echo=FALSE}
anomalies %>%
  mutate(
    response_time_ms = round(response_time_ms, 1),
    error_rate = sprintf("%.2f%%", error_rate * 100),
    cpu_usage = sprintf("%.1f%%", cpu_usage * 100)
  ) %>%
  datatable(
    options = list(
      pageLength = 10,
      dom = 'Bfrtip',
      scrollX = TRUE
    ),
    rownames = FALSE,
    class = 'cell-border stripe',
    caption = 'Operations with anomalous behavior (sorted by response time)'
  )
```

---

## Statistical Summary

```{r statistical_summary, echo=FALSE}
# Comprehensive summary statistics
summary_stats <- performance_data %>%
  summarise(
    Metric = c("Response Time (ms)", "Throughput (ops/sec)", "Error Rate (%)",
               "CPU Usage (%)", "Memory (MB)", "Concurrent Users"),
    Mean = c(mean(response_time_ms), mean(throughput_ops), mean(error_rate) * 100,
             mean(cpu_usage) * 100, mean(memory_usage_mb), mean(concurrent_users)),
    Median = c(median(response_time_ms), median(throughput_ops), median(error_rate) * 100,
               median(cpu_usage) * 100, median(memory_usage_mb), median(concurrent_users)),
    SD = c(sd(response_time_ms), sd(throughput_ops), sd(error_rate) * 100,
           sd(cpu_usage) * 100, sd(memory_usage_mb), sd(concurrent_users)),
    Min = c(min(response_time_ms), min(throughput_ops), min(error_rate) * 100,
            min(cpu_usage) * 100, min(memory_usage_mb), min(concurrent_users)),
    Max = c(max(response_time_ms), max(throughput_ops), max(error_rate) * 100,
            max(cpu_usage) * 100, max(memory_usage_mb), max(concurrent_users))
  ) %>%
  mutate(across(where(is.numeric), ~round(., 2)))

kable(summary_stats, align = 'lcccccc', caption = "Comprehensive Performance Statistics")
```

---

## Correlation Analysis

```{r correlation, echo=FALSE}
# Correlation matrix
numeric_cols <- c("response_time_ms", "throughput_ops", "error_rate",
                  "cpu_usage", "memory_usage_mb", "concurrent_users")
cor_matrix <- cor(performance_data[, numeric_cols])

# Interactive heatmap
plot_ly(
  z = cor_matrix,
  x = colnames(cor_matrix),
  y = colnames(cor_matrix),
  type = "heatmap",
  colorscale = list(c(0, "#ef4444"), c(0.5, "#fbbf24"), c(1, "#22c55e")),
  zmin = -1, zmax = 1,
  text = round(cor_matrix, 3),
  hovertemplate = '%{x} vs %{y}<br>Correlation: %{z:.3f}<extra></extra>'
) %>%
  layout(
    title = "Performance Metrics Correlation Matrix",
    xaxis = list(title = "", color = '#f1f5f9'),
    yaxis = list(title = "", color = '#f1f5f9'),
    plot_bgcolor = '#1e293b',
    paper_bgcolor = '#0f172a',
    font = list(color = '#f1f5f9')
  )
```

<div class="card">
**Key Correlations:**

```{r key_correlations, echo=FALSE, results='asis'}
# Find strongest positive and negative correlations
cor_pairs <- which(upper.tri(cor_matrix), arr.ind = TRUE)
cor_values <- cor_matrix[cor_pairs]
cor_df <- data.frame(
  var1 = rownames(cor_matrix)[cor_pairs[,1]],
  var2 = colnames(cor_matrix)[cor_pairs[,2]],
  correlation = cor_values
) %>%
  arrange(desc(abs(correlation))) %>%
  head(5)

for(i in 1:nrow(cor_df)) {
  cat(paste0("- **", cor_df$var1[i], "** vs **", cor_df$var2[i], "**: ",
             sprintf("%.3f", cor_df$correlation[i]), "\n"))
}
```
</div>

---

## Recommendations

<div class="card">

### 🎯 Performance Optimization Opportunities

1. **Response Time**: Average response time is `r round(avg_response_time, 1)`ms.
   - Investigate operations exceeding `r round(quantile(performance_data$response_time_ms, 0.95), 1)`ms (95th percentile)
   - Consider caching strategies for frequently accessed operations

2. **Error Rate**: Current error rate of `r sprintf("%.2f%%", avg_error_rate)`.
   - `r if(avg_error_rate > 2) "⚠️ ERROR RATE ABOVE 2% - IMMEDIATE ACTION REQUIRED" else "✅ Error rate within acceptable limits"`
   - Review error patterns during peak hours

3. **Resource Utilization**:
   - CPU usage averaging `r sprintf("%.1f%%", avg_cpu)`
   - Memory usage averaging `r round(mean(performance_data$memory_usage_mb), 0)`MB
   - `r if(avg_cpu > 75) "⚠️ High CPU utilization - consider scaling" else "✅ CPU utilization healthy"`

4. **Anomalies**: `r anomaly_count` anomalous operations detected (`r sprintf("%.1f%%", anomaly_pct)`)
   - Investigate root causes of performance degradation
   - Implement alerting for similar patterns

</div>

---

## About This Report

<div class="card">
**Report Details:**

- **Generated:** `r Sys.time()`
- **Analysis Period:** `r format(min(performance_data$timestamp), "%Y-%m-%d")` to `r format(max(performance_data$timestamp), "%Y-%m-%d")`
- **Total Operations:** `r format(n_records, big.mark=",")`
- **Anomalies Detected:** `r anomaly_count` (`r sprintf("%.1f%%", anomaly_pct)`)
- **Author:** Kamil Nazaruk
- **GitHub:** [r-performance-analysis](https://github.com/KamilNaz/r-performance-analysis)
- **Portfolio:** [kamilnaz.github.io](https://kamilnaz.github.io)

This report was generated using R Markdown with interactive Plotly visualizations. The analysis uses sample data for demonstration purposes. For production use, replace the sample data generation with your actual performance metrics.
</div>

```{r session_info, echo=FALSE}
cat("\n### R Session Information\n\n")
sessionInfo()
```
